---
title: "manuscript_figure_results"
output: html_document
date: "2026-01-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidymodels)
library(tidyverse)
library(ranger)
library(SuperLearner)
library(sf)
library(dplyr)
library(ggplot2)
library(WeightIt)
library(GpGp)
library(Metrics)
library(rnaturalearth)
library(ggplot2)
library(see)
library(yardstick)
library(xgboost)
library(glmnet)
library(kernlab)
library(glmnet)
library(numDeriv)
library(MASS)
library(tigris)
library(rnaturalearthdata)
library(e1071)
library(broom)
library(stringr) # For str_remove_all
library(gratia) # gam plot smooth term
library(ggeffects)
```

```{r}
source('functions.r')
```

### Load data

#### nitrate measurement

```{r}
data <- load_nitrate_data(file_path = "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_with_covar_median_well.csv", zero_inflated = TRUE)
# data <- load_nitrate_data(file_path = "/Users/xindilin/Downloads/data_Nitrate_with_covar_median_well.csv", zero_inflated = TRUE)
```

we split the data into training and validation set

```{r}
data_split <- split_nitrate_data(data)
```

This is the data_split with some the results attached

```{r}
data_split <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/data_split.rds")
```

```{r}
depth_range <- c(min(data$logWellDepth),max(data$logWellDepth))
```

### Pure kriging

```{r}
nitrate_fitted_gp <- GpGp::fit_model(data_split$data$logconcentration_plus_median,
                           locs = data_split$data[,c("longitude", "latitude")],
                           covfun_name = "matern_sphere")
saveRDS(nitrate_fitted_gp, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_gp.rds")

```

```{r}
pred_gp <- GpGp::predictions(fit = nitrate_fitted_gp, locs_pred = data_split$data_test[,c("longitude", "latitude")], X_pred = rep(1,nrow(data_split$data_test)))
```


```{r}
Metrics::rmse(data_split$data_test$logconcentration_plus_median, pred_gp)
```

```{r}
data_split$data_test$pred_krig <- pred_gp
```

### Pure Kriging with design matrix

We would ignore this, very similar to linear regression first, then kriging.

```{r}
rec <- recipe(~ ., data = data_split$data[, OR_covariate_names]) %>%
    step_dummy(all_nominal_predictors()) %>%
    prep() # This "fits" the preprocessor
  
  # 2. Apply the *same* fitted recipe to both datasets
design_matrix <- as.matrix(bake(rec, new_data = data_split$data[, OR_covariate_names]))
```


```{r}
nitrate_fitted_gp_regression <- GpGp::fit_model(data_split$data$logconcentration_plus_median,
                                                X = design_matrix,locs = data_split$data[,c("longitude", "latitude")], covfun_name = "matern_sphere")

design_matrix_test <- as.matrix(bake(rec, new_data = data_split$data_test[, OR_covariate_names]))
pred_gp_regression <- GpGp::predictions(fit = nitrate_fitted_gp_regression, locs_pred = data_split$data_test[,c("longitude", "latitude")], X_pred = design_matrix_test)
```

```{r}
Metrics::rmse(data_split$data_test$logconcentration_plus_median, pred_gp_regression)
```

### Linear Regression 

```{r}
nitrate_fitted_lm <- nitrate_prediction_SL(data = data_split$data, data_test = data_split$data_test, SL.library = "SL.lm")
```

```{r}
saveRDS(nitrate_fitted_lm, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_lm.rds")
nitrate_fitted_lm <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_lm.rds")
```

```{r}
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_lm$pred_test)
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_lm$pred_test + nitrate_fitted_lm$krige_values_test)
```

```{r}
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_lm$pred_test))
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_lm$pred_test + nitrate_fitted_lm$krige_values_test))
```

```{r}
data_split$data_test$pred_lm <- nitrate_fitted_lm$pred_test
data_split$data_test$pred_lm_krig <- nitrate_fitted_lm$pred_test + nitrate_fitted_lm$krige_values_test
```

#### scatter plot 

```{r}
plot(nitrate_fitted_lm$pred_test,data_split$data_test$logconcentration_plus_median, main = 'Linear Regression. nitrate v.s. prediction')
plot(nitrate_fitted_lm$pred_test + nitrate_fitted_lm$krige_values_test, data_split$data_test$logconcentration_plus_median, main = 'Linear Regression + kriging. nitrate v.s. prediction')
```

#### prediction nitrate map in Wisconsin

```{r}
counties <- counties(state = "WI", cb = TRUE, year = 2022) # 'cb = TRUE' for cartographic boundary (simplified) version
counties = counties[,c("NAME","geometry")]
colnames(counties) = c("County","geometry")
```

```{r}
ggplot()+
  geom_sf(data = plss_covariates_central_sf, aes(color = U), size = 0.15) +
  geom_sf(data = counties,color = 'black')+
  scale_color_viridis_c(option = "plasma",
                        name = "Estimation of U",
                        labels = function(breaks) { round(exp(breaks), 0) })+
  labs(title = "5mg/L Threshold Well Depth", x = "Longitude", y = "Latitude") +
  theme_void()
```


### ranger

This is actually very slow, so we better store the fitted model

```{r}
nitrate_fitted_ranger <- nitrate_prediction_SL(data = data_split$data, data_test = data_split$data_test, SL.library = "SL.ranger")
```

```{r}
saveRDS(nitrate_fitted_ranger, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger.rds")
```

```{r}
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_ranger$pred_test)
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_ranger$pred_test + nitrate_fitted_ranger$krige_values_test)
```


```{r}
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_ranger$pred_test))
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_ranger$pred_test + nitrate_fitted_ranger$krige_values_test))
```

```{r}
plot(nitrate_fitted_ranger$pred_test,data_split$data_test$logconcentration_plus_median, main = 'Linear Regression. nitrate v.s. prediction')
plot(nitrate_fitted_ranger$pred_test + nitrate_fitted_ranger$krige_values_test, data_split$data_test$logconcentration_plus_median, main = 'Linear Regression + kriging. nitrate v.s. prediction')
```

```{r}
data_split$data_test$pred_ranger <- nitrate_fitted_ranger$pred_test
data_split$data_test$pred_ranger_krig <- nitrate_fitted_ranger$pred_test + nitrate_fitted_ranger$krige_values_test
```

### gam

```{r}
nitrate_fitted_gam <- nitrate_prediction_gam(data = data_split$data, data_test = data_split$data_test)
```
```{r}
saveRDS(nitrate_fitted_gam, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_gam.rds")
nitrate_fitted_gam <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_gam.rds")
```

```{r}
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_gam$pred_test)
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_gam$pred_test + nitrate_fitted_gam$krige_values_test)
```

transback to original scale mg/L

```{r}
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_gam$pred_test))
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_gam$pred_test + nitrate_fitted_gam$krige_values_test))
```

```{r}
data_split$data_test$pred_gam <- nitrate_fitted_gam$pred_test
data_split$data_test$pred_gam_krig <- nitrate_fitted_gam$pred_test + nitrate_fitted_gam$krige_values_test
```

### xgboost

```{r}
nitrate_fitted_xgboost <- nitrate_prediction_xgboost(data = data_split$data, data_test = data_split$data_test)
```

```{r}
saveRDS(nitrate_fitted_xgboost, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_xgboost.rds")
nitrate_fitted_xgboost <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_xgboost.rds")
```

```{r}
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_xgboost$pred_test)
Metrics::rmse(data_split$data_test$logconcentration_plus_median, nitrate_fitted_xgboost$pred_test + nitrate_fitted_xgboost$krige_values_test)
```

Let's see the prediction error in original scale

```{r}
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_xgboost$pred_test))
Metrics::rmse(data_split$data_test$concentration_plus_median, back_trans(nitrate_fitted_xgboost$pred_test + nitrate_fitted_xgboost$krige_values_test))
```

```{r}
data_split$data_test$pred_xgboost <- nitrate_fitted_xgboost$pred_test
data_split$data_test$pred_xgboost_krig <- nitrate_fitted_xgboost$pred_test + nitrate_fitted_xgboost$krige_values_test
```


### Visualization

#### BarPlot of RMSE

```{r}
Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_gam)
```

```{r}
# 1. Create a data frame of rMSE

# with pure kriging
rmse_data <- data.frame(
  Method = c("LR", "LR", 
             "GAM", "GAM", 
             "XGBoost", "XGBoost", 
             "RF", "RF", 
             "Kriging"),
  Variant = c("Base only", "Base and Kriging", 
              "Base only", "Base and Kriging", 
              "Base only", "Base and Kriging", 
              "Base only", "Base and Kriging", 
              "Kriging Only"),
  RMSE = c(Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_lm), 
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_lm_krig),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_gam),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_gam_krig),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_xgboost),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_xgboost_krig),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_ranger),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_ranger_krig),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_krig))
)

# without pure kriging
rmse_data <- data.frame(
  Method = c("LR", "LR", 
             # "Linear Regression", "Linear Regression", 
             "GAM", "GAM", 
             # "Generalized Additive Model", "Generalized Additive Model", 
             "XGBoost", "XGBoost", 
             # "Gradient Boosting", "Gradient Boosting", 
             # "Random Forest", "Random Forest"
             "RF", "RF"),
  Variant = c("Base only", "Base and Kriging", 
              "Base only", "Base and Kriging", 
              "Base only", "Base and Kriging", 
              "Base only", "Base and Kriging"),
  RMSE = c(Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_lm), 
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_lm_krig),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_gam),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_gam_krig),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_xgboost),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_xgboost_krig),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_ranger),
           Metrics::rmse(data_split$data_test$logconcentration_plus_median, data_split$data_test$pred_ranger_krig))
)

back_trans <- function(x) exp(x) - 0.5

# transform back to the original scale
rmse_data <- data.frame(
  Method = c("LR", "LR", "GAM", "GAM", "XGBoost", "XGBoost", "RF", "RF", "Kriging"),
  Variant = c("Base only", "Base and Kriging", "Base only", "Base and Kriging", 
              "Base only", "Base and Kriging", "Base only", "Base and Kriging", "Kriging Only"),
  RMSE = c(
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_lm)), 
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_lm_krig)),
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_gam)),
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_gam_krig)),
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_xgboost)),
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_xgboost_krig)),
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_ranger)),
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_ranger_krig)),
    Metrics::rmse(back_trans(data_split$data_test$logconcentration_plus_median), back_trans(data_split$data_test$pred_krig))
  )
)
```


```{r}
# 2. Create the Barplot
ggplot(rmse_data, aes(x = reorder(Method, -RMSE), y = RMSE, fill = Variant)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = round(RMSE, 3)), 
            position = position_dodge(width = 0.8), 
            vjust = -0.5, size = 4.5) +
  scale_fill_manual(values = c("Base only" = "#A6CEE3", 
                               "Kriging Only" = "#B2DF8A",
                               "Base and Kriging" = "#1F78B4" 
                               )) +
  coord_cartesian(ylim = c(2.5,4.4))+
  theme_minimal() +
  labs(title = "Validation RMSE Comparison",
       x = "Model Type",
       y = "Root Mean Square Error (mg/L)",
       fill = "Method") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12, color = "black"),
        legend.title = element_text(size = 13, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = c(0.98, 0.98),
        legend.justification = c("right", "top"),
        legend.background = element_rect(fill = "white", color = "grey80"))
```

#### Scatterplot of Kriging v.s. No Kriging

```{r}
# Assuming 'val_data' has columns: Observed, Pred_LR, Pred_LR_Krig, etc.
val_long <- data_split$data_test %>%
  pivot_longer(
    cols = starts_with("pred_"),   # Select all your prediction columns
    names_to = "Model",            # The name of the new 'Model' column
    values_to = "Predicted"        # The name of the new 'Predicted' column
  ) %>%
  # Optional: Clean up names (e.g., change "Pred_LR_Krig" to "LR + Kriging")
  mutate(Model = gsub("pred_", "", Model),
         Model = gsub("_krig", " + kriging", Model))
# Pivot long for ggplot
ggplot(val_long, aes(x = logconcentration_plus_median, y = Predicted)) +
  geom_point(alpha = 0.4, size = 0.1) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  facet_wrap(~Model, nrow = 3) +
  theme_bw() +
  labs(title = "Goodness-of-Fit Across All 9 Models",
       x = "Measured Nitrate", y = "Predicted Nitrate")
```

#### Dumbbell Plot

```{r}
library(ggplot2)
library(dplyr)

# Sample data structure
results <- data.frame(
  Method = rep(c("Linear Regression", "GAM", "Random Forest", "XGBoost"), each = 2),
  Type   = rep(c("Base", "With Kriging"), 4),
  RMSE   = c(0.50, 0.42, 0.45, 0.38, 0.35, 0.28, 0.32, 0.25) # Replace with yours
)

# Add Pure Kriging as a reference line or separate point
ggplot(results, aes(x = RMSE, y = reorder(Method, RMSE))) +
  geom_line(aes(group = Method), color = "grey") +
  geom_point(aes(color = Type), size = 4) +
  geom_vline(xintercept = 0.40, linetype = "dashed", color = "red") + # Example: Pure Kriging RMSE
  annotate("text", x = 0.41, y = 4.5, label = "Pure Kriging", color = "red", hjust = 0) +
  theme_minimal() +
  scale_color_manual(values = c("Base" = "#E15759", "With Kriging" = "#4E79A7")) +
  labs(title = "Validation RMSE: Base Models vs. Kriging-Enhanced",
       subtitle = "The line represents the improvement gained by modeling spatial residuals",
       y = NULL, x = "RMSE (Validation Set)")
```

#### map of residuals

I was told that it is better to use log10 scale

```{r}
log10_trans <- function(x) log10(exp(x) + 0.5)
```

```{r}
# 1. Calculate Residuals for all models
# We subtract the prediction from the observed value
val_with_residuals <- data_split$data_test[,c("longitude","latitude","area","WellDepth","logconcentration_plus_median","pred_lm","pred_lm_krig","pred_gam","pred_gam_krig","pred_ranger","pred_ranger_krig","pred_xgboost","pred_xgboost_krig","pred_krig")] %>%
  mutate(
    resid_lm = logconcentration_plus_median - pred_lm,
    resid_lm_krig = logconcentration_plus_median - pred_lm_krig,
    resid_gam = logconcentration_plus_median - pred_gam,
    resid_gam_krig = logconcentration_plus_median - pred_gam_krig,
    resid_ranger = logconcentration_plus_median - pred_ranger,
    resid_ranger_krig = logconcentration_plus_median - pred_ranger_krig,
    resid_xgboost = logconcentration_plus_median - pred_xgboost,
    resid_xgboost_krig = logconcentration_plus_median - pred_xgboost_krig,
    resid_krig = logconcentration_plus_median - pred_krig
  )

#  or we can also do it on the original scale, take exp(x) + 0.5
val_with_residuals_original <- data_split$data_test %>%
  mutate(
    resid_lm           = back_trans(logconcentration_plus_median) - back_trans(pred_lm),
    resid_lm_krig      = back_trans(logconcentration_plus_median) - back_trans(pred_lm_krig),
    resid_gam          = back_trans(logconcentration_plus_median) - back_trans(pred_gam),
    resid_gam_krig     = back_trans(logconcentration_plus_median) - back_trans(pred_gam_krig),
    resid_ranger       = back_trans(logconcentration_plus_median) - back_trans(pred_ranger),
    resid_ranger_krig  = back_trans(logconcentration_plus_median) - back_trans(pred_ranger_krig),
    resid_xgboost      = back_trans(logconcentration_plus_median) - back_trans(pred_xgboost),
    resid_xgboost_krig = back_trans(logconcentration_plus_median) - back_trans(pred_xgboost_krig),
    resid_krig         = back_trans(logconcentration_plus_median) - back_trans(pred_krig)
  )

# or we can do it in a log-10 scale
val_with_residuals_log10 <- data_split$data_test %>%
  mutate(
    resid_lm           = log10_trans(logconcentration_plus_median) - log10_trans(pred_lm),
    resid_lm_krig      = log10_trans(logconcentration_plus_median) - log10_trans(pred_lm_krig),
    resid_gam          = log10_trans(logconcentration_plus_median) - log10_trans(pred_gam),
    resid_gam_krig     = log10_trans(logconcentration_plus_median) - log10_trans(pred_gam_krig),
    resid_ranger       = log10_trans(logconcentration_plus_median) - log10_trans(pred_ranger),
    resid_ranger_krig  = log10_trans(logconcentration_plus_median) - log10_trans(pred_ranger_krig),
    resid_xgboost      = log10_trans(logconcentration_plus_median) - log10_trans(pred_xgboost),
    resid_xgboost_krig = log10_trans(logconcentration_plus_median) - log10_trans(pred_xgboost_krig),
    resid_krig         = log10_trans(logconcentration_plus_median) - log10_trans(pred_krig)
  )


# 2. Pivot to Long format for faceting
# This puts all residuals in one column and model names in another
val_long <- val_with_residuals_log10 %>%
# val_long <- val_with_residuals_original %>%
# val_long <- val_with_residuals %>%
  pivot_longer(
    cols = starts_with("resid_"),
    names_to = "Model",
    values_to = "Residual"
  ) %>%
  # Optional: Clean up the model names for the plot labels
  mutate(Model = gsub("resid_", "", Model),
         Model = factor(Model, levels = c("lm", "lm_krig", "gam", "gam_krig", 
                                          "ranger", "ranger_krig", "xgboost", 
                                          "xgboost_krig", "krig")))

# 3. Convert to a spatial (sf) object
# Replace 'Longitude' and 'Latitude' with your actual coordinate column names
# crs = 4326 is standard GPS coordinates (WGS84)
val_sf <- st_as_sf(val_long, 
                   coords = c("longitude", "latitude"), 
                   crs = 4326)

val_sf <- val_sf[val_sf$Model %in% c("lm","lm_krig", "ranger","ranger_krig"),]
```

```{r}
model_names <- c(
  "lm"           = "Linear Regression",
  "lm_krig"      = "Linear Regression + Kriging",
  "ranger"       = "Random Forest",
  "ranger_krig"  = "Random Forest + Kriging"
)

# Create a residual column: resid = Observed - Predicted

ggplot() +
  # Add the county layer first (bottom layer)
  # Add your prediction points layer
  geom_sf(data = val_sf[val_sf$Model %in% c("lm","lm_krig", "ranger","ranger_krig"),], 
          aes(color = pmin(pmax(Residual,-1),1)), # trim it to make to look better
          # aes(color = Residual),
          size = 0.01) +
  geom_sf(data = counties, fill = NA, color = "grey80", size = 0.2) +
  scale_color_gradient2(low = "#000080", 
                        # mid = "grey95", 
                        mid = "#ffffbf", 
                        high = "#e31a1c", 
                        name = expression(paste("Prediction Error\n[log10(mg/L)]"))
                        # name = "Prediction Error"
                        ) +
  facet_wrap(~Model, labeller = as_labeller(model_names)) +
  theme_minimal() +
  theme(legend.position = "right",
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12))


# ggplot(val_sf[val_sf$Model %in% c("lm","lm_krig", "ranger","ranger_krig"),]) +
#   geom_sf(counties, color = "grey")+
#   geom_sf(aes(color = Residual), size = 0.02) +
#   scale_color_gradient2(low = "blue", mid = "white", high = "red",name = "Prediction Error") +
#   facet_wrap(~Model,labeller = as_labeller(model_names)) +
#   theme_void() 
# +
#   labs(title = "Spatial Distribution of Validation Residuals",
#        subtitle = "Ideal models show no spatial patterns in these maps")
```

Because there are too many overlap, I plan to divide the well depth by quantiles, and draw the residual map at different well depth range

```{r}
depth_quantile <- quantile(data_split$data_test$WellDepth)
```


```{r}
ggplot() +
  # Add the county layer first (bottom layer)
  # Add your prediction points layer
  geom_sf(data = val_sf[(val_sf$Model %in% c("lm","lm_krig", "ranger","ranger_krig")) & (val_sf$WellDepth<depth_quantile[4]) & (val_sf$WellDepth>=depth_quantile[1]),], 
          aes(color = pmin(pmax(Residual,-1),1)), # trim it to make to look better
          size = 0.01) +
  geom_sf(data = counties, fill = NA, color = "black", size = 0.05) +
  scale_color_gradient2(low = "blue", 
                        mid = "#ffffbf", 
                        high = "#e31a1c", 
                        limits = 
                        name = "Prediction Error [log10(mg/L)]") +
  facet_wrap(~Model, labeller = as_labeller(model_names)) +
  # theme_void() +
  theme_minimal() +
  theme(legend.position = "right")

```

#### 3D map

##### load PLSS covariates

generate the PLSS covariate with static water level and remove NaNs for building 3-D map

```{r}
file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/plss_covariates.csv"
command <- paste("brctl download", shQuote(file_path))
system(command)
plss_covariates <- read.csv(file_path)
```

```{r}
plss_covariates$cafolog <- log(plss_covariates$cafo + 1)
plss_covariates <- na.omit(plss_covariates)
plss_covariates <- plss_covariates[plss_covariates$crop_type_combine %in% unique(data$crop_type_combine),]
```

```{r}
gpfit_staticLevel <- GpGp::fit_model(data$StaticLevel,
                                     locs = data[,c("longitude", "latitude")],
                                     covfun_name = "matern_sphere")

plss_covariates$StaticLevel <-  GpGp::predictions(fit = gpfit_staticLevel,
                                                  locs_pred = plss_covariates[,c("longitude", "latitude")],
                                                  X_pred = rep(1,nrow(plss_covariates)))

```

```{r}
write.csv(plss_covariates, "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/plss_covariates_static_no_na.csv", row.names = TRUE)
```

```{r}
file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/plss_covariates_static_no_na.csv"
command <- paste("brctl download", shQuote(file_path))
system(command)
plss_covariates <- read.csv(file_path)
```


##### draw maps at different depths

we use the ranger predictors

```{r}
nitrate_fitted_ranger <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger.rds")
```


```{r}
library(tidyverse)
library(sf)

# 1. Define depths and combine into one master dataframe
depths <- c(50, 100, 200, 400)

all_depths_df <- map_dfr(depths, function(d) {
  plss_covariates %>%
    mutate(
      WellDepth = d,
      logWellDepth = log(d),
      Depth_Label = paste(d, "Feet") # Useful for facet titles
    )
})

# 2. Run SuperLearner predictions on the entire stack at once
all_depths_df$pred_ranger <- predict.SuperLearner(
  nitrate_fitted_ranger$SL_obj, 
  newdata = all_depths_df
)$pred

# 3. Run Kriging (GpGp)
# Note: Ensure coordinates are in matrix form
krige_output <- GpGp::predictions(
  fit = nitrate_fitted_ranger$gpfit,
  locs_pred = as.matrix(all_depths_df[, c("longitude", "latitude")]),
  X_pred = rep(1, nrow(all_depths_df))
)
all_depths_df$krige_ranger <- krige_output

# 4. Transform to original scale and apply the floor (pmax)
all_depths_df <- all_depths_df %>%
  mutate(
    pred_original = exp(pred_ranger + krige_ranger),
    pred_original = pmax(pred_original, 0.4274596)
  )

# 5. Convert to SF object once
all_depths_sf <- st_as_sf(all_depths_df, coords = c("longitude", "latitude"), crs = 4326)
```

```{r}
# plot log_10 legend
ggplot() +
  # Points Layer
  geom_sf(data = all_depths_sf, aes(color =pred_original), size = 0.01) +
  # Counties Layer (drawn on top for clarity)
  geom_sf(data = counties, fill = NA, color = "black", size = 0.1) +
  # Shared Color Scale
  scale_color_viridis_c(
    option = "plasma", 
    name = "Predicted Nitrate\n(mg/L)",
    trans = "log10", 
    # trans = "log", 
    limits = c(0.4274596, 31.52249),
    breaks = c(1, 5, 10, 20, 30),
    labels = c( "1", "5", "10", "20", "30")
  ) +
  # Create the 2x2 Grid
  facet_wrap(~reorder(Depth_Label, WellDepth), ncol = 2) +
  labs(
    title = "Nitrate Concentration Predictions by Depth",
    subtitle = "Integration of Residual Kriging and Random Forest",
    # x = "Longitude", 
    # y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "plain"),
    legend.position = "right",
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.key.width = unit(0.7, "cm")
  )
```

#### Map of Required Well Depth

It is better to use the model that is retrained with truncation, i.e., truncated below 2mg/L for 10mg/L and truncated above 10mg/L for 2mg/L threshold.

```{r}
# nitrate_fitted_ranger <- nitrate_prediction_SL(data = data_split$data, data_test = data_split$data_test, SL.library = "SL.ranger")

nitrate_fitted_ranger_less_zero <- nitrate_prediction_SL(data = data_split$data[data_split$data$Sample.Analytical.Result.Amount!=0,], 
                                                         data_test = data_split$data_test[data_split$data_test$Sample.Analytical.Result.Amount!=0,], 
                                                         SL.library = "SL.ranger")
saveRDS(nitrate_fitted_ranger_less_zero,"~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger_less_zero.rds")
nitrate_fitted_ranger_less_zero <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger_less_zero.rds")
```

```{r}
nitrate_fitted_ranger <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger.rds")
```


##### diagnosis of less_zero model

Why the previous fitted model is no different from the fitted model using all data

```{r}
nitrate_fitted_ranger_tmp <- nitrate_prediction_ranger(data = data_split$data, data_test = data_split$data_test)
nitrate_fitted_ranger_less_zero_tmp <- nitrate_prediction_ranger(data_split$data[data_split$data$Sample.Analytical.Result.Amount!=0,], 
                                                         data_test = data_split$data_test[data_split$data_test$Sample.Analytical.Result.Amount!=0,])
```


```{r}
pred_temp <- do.call(predict, c(list(object = nitrate_fitted_ranger_tmp$ranger_obj, data = nitrate_fitted_ranger_less_zero_tmp$data_test)))
pred_temp <- pred_temp$predictions
```

```{r}
plot(nitrate_fitted_ranger_less_zero_tmp$pred_test, pred_temp)
```

```{r}
pred_temp <- do.call(predict, c(list(object = nitrate_fitted_ranger$SL_obj, newdata = nitrate_fitted_ranger_less_zero$data_test)))
pred_temp <- pred_temp$pred
```

```{r}
plot(nitrate_fitted_ranger_less_zero$pred_test, pred_temp)
```

we would now use nitrate_fitted_ranger_less_two and nitrate_fitted_ranger_more_ten

```{r}
nitrate_fitted_ranger_less_two <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger_less_two.rds")
nitrate_fitted_ranger_more_ten <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger_more_ten.rds")
```

We note that nitrate_fitted_ranger_less_two, nitrate_fitted_ranger_less_two and nitrate_fitted_ranger are trained using different scale. The first two are trained on original scale, the last is trained on log-transformed scale


we load the PLSS dataset

```{r}
file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/plss_covariates_static_no_na.csv"
command <- paste("brctl download", shQuote(file_path))
system(command)
plss_covariates <- read.csv(file_path)
```

##### get the value at the depth grid at every location

```{r}
# depth_range <- c(min(data$logWellDepth),max(data$logWellDepth))
depth_range <- c(min(data$logWellDepth),quantile(data$logWellDepth,0.9997))
treatment_name <- "logWellDepth"
treatment_values <- seq(depth_range[1], depth_range[2], by = 0.05)

data_expanded <- plss_covariates[rep(seq_len(nrow(plss_covariates)), each = length(treatment_values)), ]
# the data_expanded is very large, so the following two line might be very time-consuming
data_expanded[,treatment_name] <- rep(treatment_values, times = nrow(plss_covariates))

preds_less_two <- do.call(predict, c(list(object = nitrate_fitted_ranger_less_two$SL_obj, newdata = data_expanded)))
preds_less_two <- preds_less_two$pred

preds_less_zero <- do.call(predict, c(list(object = nitrate_fitted_ranger_less_zero$SL_obj, newdata = data_expanded)))
preds_less_zero <- preds_less_zero$pred

preds_more_ten <- do.call(predict, c(list(object = nitrate_fitted_ranger_more_ten$SL_obj, newdata = data_expanded)))
preds_more_ten <- preds_more_ten$pred

preds <- do.call(predict, c(list(object = nitrate_fitted_ranger$SL_obj, newdata = data_expanded)))
preds <- preds$pred # if it is from superlearner

pred_matrix_less_two <- matrix(preds_less_two, nrow = length(treatment_values), ncol = nrow(plss_covariates))
pred_matrix_less_zero <- matrix(preds_less_zero, nrow = length(treatment_values), ncol = nrow(plss_covariates))
pred_matrix_more_ten <- matrix(preds_more_ten, nrow = length(treatment_values), ncol = nrow(plss_covariates))
pred_matrix <- matrix(preds, nrow = length(treatment_values), ncol = nrow(plss_covariates))
```

```{r}
set.seed(1998)
krige_values_PLSS_ranger <- GpGp::predictions(fit = nitrate_fitted_ranger$gpfit, 
                                              locs_pred = plss_covariates[,c("longitude", "latitude")], 
                                              X_pred = rep(1,nrow(plss_covariates)))

krige_values_PLSS_ranger_less_two <- GpGp::predictions(fit = nitrate_fitted_ranger_less_two$gpfit, 
                                              locs_pred = plss_covariates[,c("longitude", "latitude")], 
                                              X_pred = rep(1,nrow(plss_covariates)))

krige_values_PLSS_ranger_less_zero <- GpGp::predictions(fit = nitrate_fitted_ranger_less_zero$gpfit, 
                                              locs_pred = plss_covariates[,c("longitude", "latitude")], 
                                              X_pred = rep(1,nrow(plss_covariates)))

krige_values_PLSS_ranger_more_ten <- GpGp::predictions(fit = nitrate_fitted_ranger_more_ten$gpfit, 
                                              locs_pred = plss_covariates[,c("longitude", "latitude")], 
                                              X_pred = rep(1,nrow(plss_covariates)))
```

```{r}
krige_values_PLSS_required_depth_map <- list(krige_values_PLSS_ranger = krige_values_PLSS_ranger,
                                        krige_values_PLSS_ranger_less_two = krige_values_PLSS_ranger_less_two,
                                        krige_values_PLSS_ranger_less_zero = krige_values_PLSS_ranger_less_zero,
                                        krige_values_PLSS_ranger_more_ten = krige_values_PLSS_ranger_more_ten)
```

```{r}
saveRDS(krige_values_PLSS_required_depth_map, "~/Desktop/2024 summer/groundwater_pesticide/R code/krige_values_PLSS_required_depth_map.rds")
krige_values_PLSS_required_depth_map <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/krige_values_PLSS_required_depth_map.rds")
```

```{r}
krige_values_PLSS_ranger <- krige_values_PLSS_required_depth_map$krige_values_PLSS_ranger
krige_values_PLSS_ranger_less_two <- krige_values_PLSS_required_depth_map$krige_values_PLSS_ranger_less_two
krige_values_PLSS_ranger_more_ten <- krige_values_PLSS_required_depth_map$krige_values_PLSS_ranger_more_ten
krige_values_PLSS_ranger_less_zero <- krige_values_PLSS_required_depth_map$krige_values_PLSS_ranger_less_zero
```

```{r}
nitrate_prediction_at_depth_grid <- list(depth_grid = treatment_values, 
                                    plss_covariates = plss_covariates,
                                    pred_matrix_less_two = pred_matrix_less_two,
                                    pred_matrix_less_zero = pred_matrix_less_zero,
                                    pred_matrix_more_ten = pred_matrix_more_ten,
                                    pred_matrix = pred_matrix)
```

```{r}
saveRDS(nitrate_prediction_at_depth_grid, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_prediction_at_depth_grid.rds")
```

```{r}
nitrate_prediction_at_depth_grid <- readRDS("~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_prediction_at_depth_grid.rds")
plss_covariates <- nitrate_prediction_at_depth_grid$plss_covariates
treatment_values <- nitrate_prediction_at_depth_grid$depth_grid
pred_matrix_less_two <- nitrate_prediction_at_depth_grid$pred_matrix_less_two
pred_matrix_less_zero <- nitrate_prediction_at_depth_grid$pred_matrix_less_zero
pred_matrix_more_ten <- nitrate_prediction_at_depth_grid$pred_matrix_more_ten
pred_matrix <- nitrate_prediction_at_depth_grid$pred_matrix
```

Notice that the retain model less_two and more_ten is trained at the original scale

```{r}
# 1. Add the kriging values to the pred_matrix columns in one step
# sweep() adds the vector krige_values_PLSS_ranger across columns (MARGIN = 2)
adjusted_matrix <- sweep(pred_matrix, 2, krige_values_PLSS_ranger, "+")
adjusted_matrix_less_two <- sweep(pred_matrix_less_two, 2, krige_values_PLSS_ranger_less_two, "+")
adjusted_matrix_less_zero <- sweep(pred_matrix_less_zero, 2, krige_values_PLSS_ranger_less_zero, "+")
adjusted_matrix_more_ten <- sweep(pred_matrix_more_ten, 2, krige_values_PLSS_ranger_more_ten, "+")

# 2. Create a logical matrix: TRUE where condition is met
pass_condition_10mg <- adjusted_matrix < log(10.5)
pass_condition_10mg_less_two <- adjusted_matrix_less_two < 10.5 # we could use remove less two to get require depths for log(10)
pass_condition_10mg_less_zero <- adjusted_matrix_less_zero < log(10.5) # we could use remove less two to get require depths for log(10)
pass_condition_2mg <- adjusted_matrix < log(2.5)
pass_condition_2mg_more_ten <- adjusted_matrix_more_ten < 2.5 # we could use remove more ten to get require depths for log(10)

pass_condition_5mg <- adjusted_matrix < log(5.5)

# 3. Find the first TRUE index in each column
# We transpose because max.col works across rows.
# ties.method = "first" ensures we get the smallest row index (the "break" logic)

first_indices_10mg <- max.col(t(pass_condition_10mg), ties.method = "first")
first_indices_10mg_less_two <- max.col(t(pass_condition_10mg_less_two), ties.method = "first")
first_indices_10mg_less_zero <- max.col(t(pass_condition_10mg_less_zero), ties.method = "first")
first_indices_2mg <- max.col(t(pass_condition_2mg), ties.method = "first")
first_indices_2mg_more_ten <- max.col(t(pass_condition_2mg_more_ten), ties.method = "first")
first_indices_5mg <- max.col(t(pass_condition_5mg), ties.method = "first")

# 4. Handle columns where the condition was NEVER met
# If no TRUE exists in a column, max.col defaults to 1. 
# We replace those with the last row index (nrow) to match your 'if(j==dim)' logic.
has_any_pass_10mg <- colSums(pass_condition_10mg) > 0
first_indices_10mg[!has_any_pass_10mg] <- nrow(pred_matrix)

has_any_pass_10mg_less_two <- colSums(pass_condition_10mg_less_two) > 0
first_indices_10mg_less_two[!has_any_pass_10mg_less_two] <- nrow(pred_matrix)

has_any_pass_10mg_less_zero <- colSums(pass_condition_10mg_less_zero) > 0
first_indices_10mg_less_zero[!has_any_pass_10mg_less_zero] <- nrow(pred_matrix_less_zero)

has_any_pass_2mg <- colSums(pass_condition_2mg) > 0
first_indices_2mg[!has_any_pass_2mg] <- nrow(pred_matrix)

has_any_pass_2mg <- colSums(pass_condition_2mg_more_ten) > 0
first_indices_2mg_more_ten[!has_any_pass_2mg] <- nrow(pred_matrix)

has_any_pass_5mg <- colSums(pass_condition_5mg) > 0
first_indices_5mg[!has_any_pass_5mg] <- nrow(pred_matrix)

# 5. Extract the actual treatment values
indirect_policy_ranger_10mg <- treatment_values[first_indices_10mg]
indirect_policy_ranger_10mg_less_zero <- treatment_values[first_indices_10mg_less_zero]
indirect_policy_ranger_10mg_less_two <- treatment_values[first_indices_10mg_less_two]


indirect_policy_ranger_2mg <- treatment_values[first_indices_2mg]
indirect_policy_ranger_2mg_more_ten <- treatment_values[first_indices_2mg_more_ten]
indirect_policy_ranger_5mg <- treatment_values[first_indices_5mg]
```

```{r}
plss_covariates$indirect_policy_ranger_10mg <- indirect_policy_ranger_10mg
plss_covariates$indirect_policy_ranger_10mg_less_two <- indirect_policy_ranger_10mg_less_two
plss_covariates$indirect_policy_ranger_10mg_less_zero <- indirect_policy_ranger_10mg_less_zero
plss_covariates$indirect_policy_ranger_2mg <- indirect_policy_ranger_2mg
plss_covariates$indirect_policy_ranger_2mg_more_ten <- indirect_policy_ranger_2mg_more_ten
plss_covariates$indirect_policy_ranger_5mg <- indirect_policy_ranger_5mg
plss_covariates_sf <- st_as_sf(plss_covariates, coords = c('longitude','latitude'), crs= 4326)
```


```{r}
counties <- counties(state = "WI", cb = TRUE, year = 2022) # 'cb = TRUE' for cartographic boundary (simplified) version
counties = counties[,c("NAME","geometry")]
colnames(counties) = c("County","geometry")
```

##### diagnosis for 10mg/L required depth deeper than 2mg/L required depth

```{r}
sum(indirect_policy_ranger_10mg>indirect_policy_ranger_2mg)
sum(indirect_policy_ranger_10mg_less_two>indirect_policy_ranger_2mg)
sum(indirect_policy_ranger_10mg_less_zero>indirect_policy_ranger_2mg)
```

```{r}
max(indirect_policy_ranger_10mg)
max(indirect_policy_ranger_10mg_less_two)
max(indirect_policy_ranger_2mg)
quantile(data$logWellDepth,0.9997)
```

```{r}
index <- indirect_policy_ranger_10mg_less_two>indirect_policy_ranger_2mg
```


```{r}
ggplot() +
  geom_sf(data = counties,color = 'black') +
  geom_sf(data = plss_covariates_sf[index,], aes(color = pmin(indirect_policy_ranger_10mg_less_two,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  scale_color_viridis_c(option = "H",
                        name = "Feet",
                        trans = "log",
                        limits = c(log(10),log(1350)),
                        breaks = log(c(10, 40, 200,1000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "10 mg/L Threshold Diagnosis: Required Depth 10mg/L Threshold > 2mg/L Threshold", x = "Longitude", y = "Latitude") +
  theme_minimal()
ggplot() +
  geom_sf(data = counties,color = 'black') +
  geom_sf(data = plss_covariates_sf[index,], aes(color = pmin(indirect_policy_ranger_2mg,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  scale_color_viridis_c(option = "H",
                        name = "Feet",
                        trans = "log",
                        limits = c(log(10),log(1350)),
                        breaks = log(c(10, 40, 200,1000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "2 mg/L Threshold Diagnosis: Required Depth 10mg/L Threshold > 2mg/L Threshold", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

##### draw required depth maps

Since we train the model using different data, i.e., truncating <2mg/L obs for required map against 10mg/L, using all obs for required map against 2mg/L.

```{r}
ggplot() +
  geom_sf(data = plss_covariates_sf, aes(color = pmin(indirect_policy_ranger_2mg_more_ten,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  # geom_sf(data = counties,color = 'black') +
  scale_color_viridis_c(option = "H",
                        name = "Feet",
                        trans = "log",
                        limits = c(log(10),log(1350)),
                        breaks = log(c(10, 40, 200,1000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  # scale_colour_gradient(low = "white",
  #                       high = "darkblue",
  #                       name = "Feet",
  #                       limits = c(log(10),log(1350)),
  #                       breaks = log(c(10, 40, 200,1000)),
  #                       labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "2mg/L Threshold Well Depth Truncate 10mg/L", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
ggplot() +
  geom_sf(data = counties,color = 'black') +
  geom_sf(data = plss_covariates_sf, aes(color = pmin(indirect_policy_ranger_2mg,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  scale_color_viridis_c(option = "H",
                        name = "Feet",
                        trans = "log",
                        limits = c(log(10),log(1350)),
                        breaks = log(c(10, 40, 200,1000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  # scale_colour_gradient(low = "white",
  #                       high = "darkblue",
  #                       name = "Feet",
  #                       limits = c(log(10),log(1350)),
  #                       breaks = log(c(10, 40, 200,1000)),
  #                       labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "2mg/L Threshold Well Depth Full Dataset", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
ggplot() +
  # geom_sf(data = counties,color = 'black') +
  geom_sf(data = plss_covariates_sf, aes(color = pmin(indirect_policy_ranger_5mg,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  scale_color_viridis_c(option = "D",
                        name = "Feet",
                        limits = c(log(7),log(2000)),
                        breaks = log(c(10, 100, 500,1000,2000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "5mg/L Threshold Well Depth", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
ggplot() +
  # geom_sf(data = counties,color = 'black') +
  geom_sf(data = plss_covariates_sf, aes(color = pmin(indirect_policy_ranger_10mg_less_zero,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  scale_color_viridis_c(option = "H",
                        name = "Feet",
                        limits = c(log(10),log(1350)),
                        breaks = log(c(10, 40, 200,1000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "10mg/L Threshold Well Depth Truncate at 0 mg/L", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
ggplot() +
  # geom_sf(data = counties,color = 'black') +
  geom_sf(data = plss_covariates_sf, aes(color = pmin(indirect_policy_ranger_10mg,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  scale_color_viridis_c(option = "H",
                        name = "Feet",
                        limits = c(log(10),log(1350)),
                        breaks = log(c(10, 40, 200,1000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "10mg/L Threshold Well Depth Full Dataset", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
ggplot() +
  geom_sf(data = plss_covariates_sf, aes(color = pmin(indirect_policy_ranger_10mg_less_two,quantile(data$logWellDepth,0.9997))), size = 0.1) +
  # geom_sf(data = counties,color = 'black') +
  scale_color_viridis_c(option = "H",
                        # option = "D",
                        name = "Feet",
                        trans = "log",
                        limits = c(log(10),log(1350)),
                        breaks = log(c(10, 40, 200,1000)),
                        labels = function(breaks) {round(exp(breaks), 0)}) +
  # scale_colour_gradient(low = "white",
  #                       high = "darkblue",
  #                       name = "Feet",
  #                       limits = c(log(10),log(1350)),
  #                       breaks = log(c(10, 40, 200,1000)),
  #                       # limits = c(log(10),log(400)),
  #                       # breaks = log(c(10, 20, 100,400)),
  #                       labels = function(breaks) {round(exp(breaks), 0)}) +
  labs(title = "10mg/L Threshold Well Depth Truncate 2mg/L", x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
depth_cap <- quantile(data$logWellDepth, 0.9997, na.rm = TRUE)
plss_long_sf <- plss_covariates %>%
  dplyr::select(longitude, latitude, indirect_policy_ranger_2mg, indirect_policy_ranger_10mg_less_zero) %>%
  pivot_longer(cols = starts_with("indirect_policy"), 
               names_to = "Threshold", 
               values_to = "Depth") %>%
  mutate(
    # Apply the cap and rename for the plot titles
    Depth_Capped = pmin(Depth, depth_cap),
    Threshold = recode(Threshold, 
                       "indirect_policy_ranger_2mg" = "2mg/L Threshold", 
                       "indirect_policy_ranger_10mg_less_zero" = "10mg/L Threshold")
  ) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
```

```{r}
meters <- c(5, 30, 80, 300)

ggplot() +
  # 1. Background layer
  geom_sf(data = counties, fill = NA, color = "grey80", size = 0.1) +
  
  # 2. Prediction points
  geom_sf(data = plss_long_sf, aes(color = Depth_Capped), size = 0.05, alpha = 0.6) +
  
  # 3. Shared Color Scale
  # we will choose 5m, 50m, 100m ,300m
  scale_color_viridis_c(
    option = "turbo", # "H" in newer ggplot is turbo
    name = "Required Well Depth\n(Meter)",
    trans = "log",
    limits = c(log(10), log(1350)),
    # breaks = log(c(10, 40, 200, 1000)),
    breaks = log(meters/0.3048),
    labels = function(x) round(exp(x)*0.3048, 0)
  ) +
  
  # 4. Faceting into 1 row, 2 columns
  facet_wrap(~Threshold, ncol = 2) +
  
  # 5. Styling
  labs(
    title = "Minimum Safe Well Depth for Nitrate Thresholds",
    subtitle = "Comparison of 2mg/L and 10mg/L regulatory targets",
    x = NULL, y = NULL
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 14, face = "plain"), # Facet Titles
    legend.position = "right",
    legend.key.width = unit(0.5, "cm"),
    panel.spacing = unit(2, "lines") # Adds space between the two maps
    # axis.text = element_blank(),      # Removes lat/long numbers for a cleaner look
    # axis.ticks = element_blank()
  )
```

#### Binary Classification

```{r}
# Load necessary libraries
# install.packages(c("pROC", "caret", "MLmetrics", "ggplot2"))
library(pROC)
library(caret)
library(MLmetrics)
library(ggplot2)
library(yardstick)
```

We can use 10mg/L or 5mg/L as threshold.

Based our discussion with Chris on Feb 24. We would use 2mg/L rather than 5mg/L as the threshold.

We found that the sensitivity is very low, this might because we include too much zero in it, i.e., the data is imbalanced.

We remove threshold nitrate levels lower than 2mg/L and retrain the model, because that is a natural nitrate level unaffected by human activities

We found that even if we remove them, it is sensitivity when using 10mg/L as threshold is still very very low, 0.3179697, 0.2170362, 0.0000000, 0.2567095, we use use the original scale to train, and see if this can be improved

##### retrain the model

Due to the data imbalance, we remove the obs lower than 2mg/L to do the binary classification

```{r}
data_less_two <- data[data$concentration_plus_median>=2,]
data_split_less_two <- split_nitrate_data(data_less_two)
```

we should notice that the response_name is ln by default, we specificy it as the original scale in the following code

```{r}
nitrate_fitted_ranger_less_two <- nitrate_prediction_SL(data = data_split_less_two$data, data_test = data_split_less_two$data_test, SL.library = "SL.ranger", response_name = "concentration_plus_median")
nitrate_fitted_lm_less_two <- nitrate_prediction_SL(data = data_split_less_two$data, data_test = data_split_less_two$data_test, SL.library = "SL.lm",response_name = "concentration_plus_median")
```

```{r}
saveRDS(nitrate_fitted_ranger_less_two, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger_less_two.rds")
saveRDS(nitrate_fitted_lm_less_two, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_lm_less_two.rds")
```


```{r}
data_split_less_two$data_test$pred_lm <- nitrate_fitted_lm_less_two$pred_test
data_split_less_two$data_test$pred_lm_krig <- nitrate_fitted_lm_less_two$pred_test + nitrate_fitted_lm_less_two$krige_values_test
```

```{r}
data_split_less_two$data_test$pred_ranger <- nitrate_fitted_ranger_less_two$pred_test
data_split_less_two$data_test$pred_ranger_krig <- nitrate_fitted_ranger_less_two$pred_test + nitrate_fitted_ranger_less_two$krige_values_test
```

##### Draw plot for 10mg/L and 5mg/L thereshold with less than two dataset

```{r}
# 1. Configuration
# Internal variable names
internal_names <- c("pred_ranger_krig", "pred_ranger", "pred_lm", "pred_lm_krig")

# Desired display names (mapped 1-to-1)
display_names <- c("RF + Kriging", "RF", 
                   "LR", "LR + Kriging")

# Create a lookup named vector
name_map <- setNames(display_names, internal_names)
```


```{r}
library(caret)
library(pROC)
library(yardstick)
library(MLmetrics)
library(ggplot2)
library(dplyr)
library(tidyr)

# 1. Configuration
pred_names <- c("pred_ranger_krig", "pred_ranger", "pred_lm", "pred_lm_krig")
threshold <- 5
test_data <- data_split_less_two$data_test

# Initialize storage
all_metrics <- data.frame()
roc_list <- list()
conf_mat_list <- list()

combined_cm_data <- data.frame()

# 2. Loop through each model to calculate metrics
for (threshold in c(5,10)){
  for (m in pred_names) {
    print(threshold)
    
    # Binary outcomes
    y_obs <- factor(ifelse(test_data$concentration_plus_median > threshold, "High", "Low"), 
                    levels = c("High", "Low"))
    y_pred <- factor(ifelse(test_data[[m]] > threshold, "High", "Low"), 
                     levels = c("High", "Low"))
    
    # Confusion Matrix metrics
    conf_mat <- confusionMatrix(y_pred, y_obs, positive = "High")
    
    df <- as.data.frame(conf_mat$table)
    df$Method <- name_map[m]
    df$Threshold_Level <- paste(threshold,"mg/L")
    
    combined_cm_data <- rbind(combined_cm_data, df)
    
    # conf_mat_list[[m]] <- conf_mat
    
    
    # MCC and F1
    eval_df <- data.frame(truth = y_obs, estimate = y_pred)
    mcc_val <- mcc(eval_df, truth = truth, estimate = estimate)$.estimate
    f1_val  <- F1_Score(y_obs, y_pred)
    
    # ROC and AUC
    # Use the continuous predictions (test_data[[m]]) for the ROC curve
    roc_obj <- roc(test_data$concentration_plus_median > threshold, 
                   test_data[[m]], quiet = TRUE)
    auc_val <- as.numeric(auc(roc_obj))
    roc_list[[m]] <- roc_obj # Save for plotting
    
    # Combine into dataframe
    model_stats <- data.frame(
      Method = name_map[m],
      Accuracy = conf_mat$overall["Accuracy"],
      Sensitivity = conf_mat$byClass["Sensitivity"],
      Specificity = conf_mat$byClass["Specificity"],
      MCC = mcc_val,
      F1 = f1_val,
      AUC = auc_val,
      Threshold_Level = threshold
    )
    all_metrics <- rbind(all_metrics, model_stats)
  }
}


# all_metrics$Method = c("Random Forest + Kriging", "Random Forest", "Linear Regression", "Linear Regression + Kriging")

# 3. Visualization: Metric Comparison Plot
plot_data <- all_metrics %>%
  pivot_longer(cols = -Method, names_to = "Metric", values_to = "Score")

ggplot(plot_data, aes(x = Method, y = Score, fill = Method)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~Metric, scales = "free_y") +
  theme_minimal() +
  labs(title = "Comparative Model Performance", subtitle = paste("Threshold:", round(threshold, 2), "mg/L")) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_fill_brewer(palette = "Set1")

# 4. Visualization: Combined ROC Curves
# # Convert list of ROC objects to a format pROC can plot together
# plot(roc_list[[1]], col = "#e41a1c", lwd = 2, main = "Combined ROC Curve Comparison")
# plot(roc_list[[2]], add = TRUE, col = "#377eb8", lwd = 2)
# plot(roc_list[[3]], add = TRUE, col = "#4daf4a", lwd = 2)
# plot(roc_list[[4]], add = TRUE, col = "#ff7f00", lwd = 2)
# legend("bottomright", legend = pred_names, 
#        col = c("#e41a1c", "#377eb8", "#4daf4a", "#ff7f00"), lwd = 2, cex = 0.8)

plot(roc_list[[1]], 
     col = "#e41a1c", 
     lwd = 2, 
     legacy.axes = TRUE,     # Changes x-axis to 1-Specificity (0 to 1)
     asp = NA,               # Prevents the plot from being forced into a perfect square
     main = paste("ROC Curves at",round(threshold, 2), "mg/L", "Threshold"))

# Add the other lines as before
plot(roc_list[[2]], add = TRUE, col = "#377eb8", lwd = 2)
plot(roc_list[[3]], add = TRUE, col = "#4daf4a", lwd = 2)
plot(roc_list[[4]], add = TRUE, col = "#ff7f00", lwd = 2)

legend("bottomright", legend = c("Random Forest + Kriging", "Random Forest", "Linear Regression", "Linear Regression + Kriging"), 
       col = c("#e41a1c", "#377eb8", "#4daf4a", "#ff7f00"), lwd = 2, cex = 0.8)
```

```{r}

names(conf_mat_list) <- c("Random Forest + Kriging", "Random Forest", "Linear Regression", "Linear Regression + Kriging")
all_cm_data <- data.frame()

for (m in names(conf_mat_list)) {
  # Convert the table to a data frame
  df <- as.data.frame(conf_mat_list[[m]]$table)
  df$Method <- m
  all_cm_data <- rbind(all_cm_data, df)
}
```

```{r}
ggplot(all_cm_data, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile(color = "white") +
  # Add the counts in the center of the tiles
  geom_text(aes(label = Freq), size = 5, fontface = "bold") +
  # Use a color scale that highlights the "High" counts
  scale_fill_gradient(low = "#f7fbff", high = "#084594") +
  # Facet by Method to create the 2x2 or 1x4 grid
  facet_wrap(~Method) +
  theme_minimal() +
  labs(
    title = "Confusion Matrix Comparison",
    subtitle = "Threshold: 10 mg/L (EPA MCL)",
    x = "Predicted Nitrate Level",
    y = "Observed Nitrate Level",
    fill = "Count"
  ) +
  theme(
    strip.text = element_text(size = 12, face = "bold"), # Headers for each plot
    panel.grid = element_blank()
  )
```

```{r}

c("Random Forest + Kriging", "Random Forest", "Linear Regression", "Linear Regression + Kriging")
```


```{r}
ggplot(combined_cm_data, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile(color = "white") +
  # Add the counts in the center
  geom_text(aes(label = Freq), size = 4, fontface = "bold") +
  # Use a consistent color scale
  scale_fill_gradient(low = "#f7fbff", high = "#084594") +
  # Create the 4 (Methods) x 2 (Thresholds) Grid
  facet_grid(Threshold_Level ~ Method) +
  theme_minimal() +
  labs(
    title = "Confusion Matrix Comparison across Methods and Thresholds",
    x = "Predicted Nitrate Level",
    y = "Observed Nitrate Level",
    fill = "Count"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank(),
    panel.border = element_rect(color = "gray90", fill = NA)
  )
```

##### draw plot for 2mg/L (and potentially 10mg/L) thereshold with original dataset

I think this does not work very well. 

##### Retrain Model (Remove above 10mg/L)

We can potentially remove obs higher than 10mg/L to train the 2mg/L model

```{r}
data_more_ten <- data[data$concentration_plus_median<=10,]
data_split_more_ten <- split_nitrate_data(data_more_ten)
```

```{r}
nitrate_fitted_ranger_more_ten <- nitrate_prediction_SL(data = data_split_more_ten$data, data_test = data_split_more_ten$data_test, SL.library = "SL.ranger", response_name = "concentration_plus_median")
nitrate_fitted_lm_more_ten <- nitrate_prediction_SL(data = data_split_more_ten$data, data_test = data_split_more_ten$data_test, SL.library = "SL.lm",response_name = "concentration_plus_median")
```
```{r}
saveRDS(nitrate_fitted_ranger_more_ten, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_ranger_more_ten.rds")

saveRDS(nitrate_fitted_lm_more_ten, "~/Desktop/2024 summer/groundwater_pesticide/R code/nitrate_fitted_lm_more_ten.rds")
```

```{r}
data_split_more_ten$data_test$pred_lm <- nitrate_fitted_lm_more_ten$pred_test
data_split_more_ten$data_test$pred_lm_krig <- nitrate_fitted_lm_more_ten$pred_test + nitrate_fitted_lm_more_ten$krige_values_test
```

```{r}
data_split_more_ten$data_test$pred_ranger <- nitrate_fitted_ranger_more_ten$pred_test
data_split_more_ten$data_test$pred_ranger_krig <- nitrate_fitted_ranger_more_ten$pred_test + nitrate_fitted_ranger_more_ten$krige_values_test
```

```{r}
# 1. Configuration
# Internal variable names
internal_names <- c("pred_ranger_krig", "pred_ranger", "pred_lm", "pred_lm_krig")

# Desired display names (mapped 1-to-1)
display_names <- c("RF + Kriging", "RF", 
                   "LR", "LR + Kriging")

# Create a lookup named vector
name_map <- setNames(display_names, internal_names)
```

```{r}
library(caret)
library(pROC)
library(yardstick)
library(MLmetrics)
library(ggplot2)
library(dplyr)
library(tidyr)
```


```{r}
data_split$data_test$pred_lm <- nitrate_fitted_lm$pred_test
data_split$data_test$pred_lm_krig <- nitrate_fitted_lm$pred_test + nitrate_fitted_lm$krige_values_test
```

```{r}
data_split$data_test$pred_ranger <- nitrate_fitted_ranger$pred_test
data_split$data_test$pred_ranger_krig <- nitrate_fitted_ranger$pred_test + nitrate_fitted_ranger$krige_values_test
```

```{r}
# 1. Configuration
pred_names <- c("pred_ranger_krig", "pred_ranger", "pred_lm", "pred_lm_krig")
test_data <- data_split$data_test
# test_data <- data_split_more_ten$data_test

# Initialize storage
all_metrics <- data.frame()
roc_list <- list()
conf_mat_list <- list()

combined_cm_data <- data.frame()

# 2. Loop through each model to calculate metrics
for (threshold in c(2)){
  for (m in pred_names) {
    print(threshold)
    
    # Binary outcomes
    # y_obs <- factor(ifelse(test_data$concentration_plus_median > threshold, "High", "Low"), 
    #                 levels = c("High", "Low"))
    # 
    # y_pred <- factor(ifelse(test_data[[m]] > threshold, "High", "Low"), 
    #                  levels = c("High", "Low"))
    
    ## use this if it is potentially trained and predict at log scale
    y_obs <- factor(ifelse(test_data$concentration_plus_median > log(threshold), "High", "Low"), 
                    levels = c("High", "Low"))
    
    y_pred <- factor(ifelse(test_data[[m]] > log(threshold), "High", "Low"), 
                     levels = c("High", "Low"))
    
    # Confusion Matrix metrics
    conf_mat <- confusionMatrix(y_pred, y_obs, positive = "High")
    
    df <- as.data.frame(conf_mat$table)
    df$Method <- name_map[m]
    df$Threshold_Level <- paste(threshold,"mg/L")
    
    combined_cm_data <- rbind(combined_cm_data, df)
    
    # conf_mat_list[[m]] <- conf_mat
    
    
    # MCC and F1
    eval_df <- data.frame(truth = y_obs, estimate = y_pred)
    mcc_val <- mcc(eval_df, truth = truth, estimate = estimate)$.estimate
    f1_val  <- F1_Score(y_obs, y_pred)
    
    # ROC and AUC
    # Use the continuous predictions (test_data[[m]]) for the ROC curve
    roc_obj <- roc(test_data$concentration_plus_median > threshold, 
                   test_data[[m]], quiet = TRUE)
    auc_val <- as.numeric(auc(roc_obj))
    roc_list[[m]] <- roc_obj # Save for plotting
    
    # Combine into dataframe
    model_stats <- data.frame(
      Method = name_map[m],
      Accuracy = conf_mat$overall["Accuracy"],
      Sensitivity = conf_mat$byClass["Sensitivity"],
      Specificity = conf_mat$byClass["Specificity"],
      MCC = mcc_val,
      F1 = f1_val,
      AUC = auc_val,
      Threshold_Level = threshold
    )
    all_metrics <- rbind(all_metrics, model_stats)
  }
}

# set Threshold_Level as NULL if there is only one threshold
all_metrics$Threshold_Level <- NULL
# all_metrics$Method = c("Random Forest + Kriging", "Random Forest", "Linear Regression", "Linear Regression + Kriging")

# 3. Visualization: Metric Comparison Plot
plot_data <- all_metrics %>%
  pivot_longer(cols = -Method, names_to = "Metric", values_to = "Score")

ggplot(plot_data, aes(x = Method, y = Score, fill = Method)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~Metric, scales = "free_y") +
  theme_minimal() +
  labs(title = "Comparative Model Performance", subtitle = paste("Threshold:", round(threshold, 2), "mg/L")) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_fill_brewer(palette = "Set1")

# 4. Visualization: Combined ROC Curves
# # Convert list of ROC objects to a format pROC can plot together
# plot(roc_list[[1]], col = "#e41a1c", lwd = 2, main = "Combined ROC Curve Comparison")
# plot(roc_list[[2]], add = TRUE, col = "#377eb8", lwd = 2)
# plot(roc_list[[3]], add = TRUE, col = "#4daf4a", lwd = 2)
# plot(roc_list[[4]], add = TRUE, col = "#ff7f00", lwd = 2)
# legend("bottomright", legend = pred_names, 
#        col = c("#e41a1c", "#377eb8", "#4daf4a", "#ff7f00"), lwd = 2, cex = 0.8)

plot(roc_list[[1]], 
     col = "#e41a1c", 
     lwd = 2, 
     legacy.axes = TRUE,     # Changes x-axis to 1-Specificity (0 to 1)
     asp = NA,               # Prevents the plot from being forced into a perfect square
     main = paste("ROC Curves at",round(threshold, 2), "mg/L", "Threshold"))

# Add the other lines as before
plot(roc_list[[2]], add = TRUE, col = "#377eb8", lwd = 2)
plot(roc_list[[3]], add = TRUE, col = "#4daf4a", lwd = 2)
plot(roc_list[[4]], add = TRUE, col = "#ff7f00", lwd = 2)

legend("bottomright", legend = c("Random Forest + Kriging", "Random Forest", "Linear Regression", "Linear Regression + Kriging"), 
       col = c("#e41a1c", "#377eb8", "#4daf4a", "#ff7f00"), lwd = 2, cex = 0.8)
```



```{r}

c("Random Forest + Kriging", "Random Forest", "Linear Regression", "Linear Regression + Kriging")
```


```{r}
ggplot(combined_cm_data, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile(color = "white") +
  # Add the counts in the center
  geom_text(aes(label = Freq), size = 4, fontface = "bold") +
  # Use a consistent color scale
  scale_fill_gradient(low = "#f7fbff", high = "#084594") +
  # Create the 4 (Methods) x 2 (Thresholds) Grid
  facet_grid(Threshold_Level ~ Method) +
  theme_minimal() +
  labs(
    title = "Confusion Matrix Comparison Without Nitrate Truncation",
    x = "Predicted Nitrate Level",
    y = "Observed Nitrate Level",
    fill = "Count"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank(),
    panel.border = element_rect(color = "gray90", fill = NA)
  )
```

### Visualize Data

```{r}
data <- load_nitrate_data(file_path = "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_with_covar_median_well.csv", zero_inflated = TRUE)
```

```{r}
file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/plss_covariates_static_no_na.csv"
command <- paste("brctl download", shQuote(file_path))
system(command)
plss_covariates <- read.csv(file_path)
plss_covariates_sf <- st_as_sf(plss_covariates, crs = 4326,coords = c("longitude","latitude"))
```

```{r}
library(ggplot2)
library(rnaturalearth)
```

```{r}
usa_states <- ne_states(country = "united states of america", returnclass = "sf")
wisconsin <- usa_states[usa_states$name == "Wisconsin", ]
```

##### CAFO plot

```{r}
ggplot() +
  geom_sf(data = wisconsin, fill = "lightgray", color = "black") +  # Wisconsin boundary
  geom_sf(data = plss_covariates_sf, aes(color= log(cafo+1)),size=0.01) +
  scale_color_viridis_c(option = "magma",name = "Ln(CAFO+1)") +  # Color scale
  theme_minimal()
```
##### Land Use Plot

```{r}
# land use/crop type
ggplot() +
  geom_sf(data = wisconsin, fill = "white", color = "white") +  # Wisconsin boundary
  geom_sf(data = plss_covariates_sf[!(plss_covariates_sf$crop_type_combine %in% c("Undefined")),], aes(color=crop_type_combine ),size=0.01) +
  scale_color_manual(
    values = c(
      "Forest" = "darkgreen", 
      "Grass" = "lightgreen", 
      "Corn" = "red",  # Make "Corn" stand out with red
      "Developed" = "gray", 
      "Soybeans" = "blue",  # Make "Soybeans" stand out with blue
      "Other Crop" = "orange",  # Make "Other Crop" stand out with orange
      "Cover Crop" = "brown", 
      "Vegtables" = "purple"
    ),
    name = "Land Use"
  ) + 
  # scale_color_viridis_d(option = "plasma", name = "Crop Type") +  # Color scale
  # labs(title = "Log(Cafo)", x = "Longitude", y = "Latitude") +
  theme_minimal() +
  guides(
    color = guide_legend(
      keywidth = unit(0.7, "cm"),  # Increase width of the legend squares
      keyheight = unit(0.7, "cm"),  # Increase height of the legend squares
      override.aes = list(shape = 15, size = 7)  # Set legend keys to squares (shape 15) and increase size
    )
  )
```

```{r}
table(plss_covariates_sf$crop_type_combine)
```

##### Precipitation Plot

```{r}
ggplot() +
  geom_sf(data = wisconsin, fill = "lightgray", color = "black") +  # Wisconsin boundary
  geom_sf(data = plss_covariates_sf, aes(color= precipitation),size=0.01) +
  scale_color_gradient(low = 'white',  # Light blue
  high = "skyblue", # Dark blue
  name = "mm/day")+
  theme_minimal()
```

##### Soil Drainage Plot

```{r}
data_factor <- factor(data,
  levels = c("Excessively drained", "Well drained", "Moderately well drained", "Somewhat poorly drained", "Poorly drained", "Very poorly drained")
)
```


```{r}

plss_covariates_sf_drainage = plss_covariates_sf[!is.na(plss_covariates_sf$drainagecl),]
plss_covariates_sf_drainage$drainagecl = factor(plss_covariates_sf_drainage$drainagecl,
  levels = c("Very poorly drained","Poorly drained","Somewhat poorly drained","Moderately well drained","Well drained","Somewhat excessively drained","Excessively drained"))

ggplot() +
  geom_sf(data = wisconsin, fill = "white", color = "white") +  # Wisconsin boundary
  geom_sf(data = plss_covariates_sf_drainage, aes(color=drainagecl),size=0.01) +
  scale_color_brewer(palette = "magma", name = "Soil Drainage Level",
                     labels = c(
                          "Very poorly drained" = "Very poor",  # Custom label
                          "Poorly drained" = "Poorly",                  # Custom label
                          "Somewhat poorly drained" = "Somewhat poor",              # Custom label
                          "Moderately well drained" = "Moderately well",         # Custom label
                          "Well drained" = "Well",
                          "Somewhat excessively drained" = "Somewhat excessive",
                          "Excessively drained" = "Excessive"
    )) + 
  theme_minimal()+
  guides(
    color = guide_legend(
      keywidth = unit(0.7, "cm"),  # Increase width of the legend squares
      keyheight = unit(0.7, "cm"),  # Increase height of the legend squares
      override.aes = list(shape = 15, size = 7)  # Set legend keys to squares (shape 15) and increase size
    )
  )
```

##### nitrate 

```{r}
ggplot() +
  geom_sf(data = wisconsin, fill = "lightgray", color = "black") +  # Wisconsin boundary
  geom_sf(data = data_sf, aes(color= log(concentration_plus_median)),size=0.01) +
  scale_color_viridis_c(option = "plasma", name = "mg/L", labels = trans_format("exp",label_number()),breaks = log(c(1,3,7, 20,50))) +  # Color scale
  theme_minimal()
```